---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(nanosvgr)
library(dplyr)
library(ggplot2)
library(tibble)

if (FALSE) {
# Check for functions which do not have an @examples block for roxygen
system("grep -c examples man/*Rd", intern = TRUE) |> 
  grep(":0$", x = _, value = TRUE)
}

if (FALSE) {
  covr::report(covr::package_coverage(
    line_exclusions = list()
  ))
}

if (FALSE) {
  pkgdown::build_site(override = list(destination = "../coolbutuseless.github.io/package/nanosvgr"))
}

# Makevars options to do some deep testing for CRAN

# Type conversions are sane
# PKG_FLAG=-Wconversion

# Pointer overflow checks i.e. dodgy pointer arithmetic
# PKG_CFLAGS+=-fsanitize=pointer-overflow -fsanitize-trap=pointer-overflow
# Then run in the debugger:
# R -d lldb 
# run
# testthat::test_local()

```

# nanosvgr

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R-CMD-check](https://github.com/coolbutuseless/nanosvgr-dev/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/nanosvgr-dev/actions/workflows/R-CMD-check.yaml)
<!-- [![CRAN status](https://www.r-pkg.org/badges/version/nanosvgr)](https://CRAN.R-project.org/package=nanosvgr) -->
<!-- badges: end -->

`nanosvgr` reads SVG  images as geometry constructed from cubic beziers
using the [`nanosvg` C library](https://github.com/memononen/nanosvg).

The beziers are further processed into polylines for easier manipulation
and plotting.

Possible future features

* Implement bezier-to-lines conversion in C.
* Add support for radial and linear gradients in `nsvg_to_grob()`. I'm unsure
  that `nanosvg` library returns enough information to do this properly.
  
  
Limitations:

* `plot.nsvg()` does not respect radial/linear gradients
* No support for SVG `<text>` in `nanosvg` library


## Installation

<!-- This package can be installed from CRAN -->

<!-- ``` r -->
<!-- install.packages('nanosvgr') -->
<!-- ``` -->

You can install the latest development version from
[GitHub](https://github.com/coolbutuseless/nanosvgr) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/nanosvgr')
```

<!-- Pre-built source/binary versions can also be installed from -->
<!-- [R-universe](https://r-universe.dev) -->

<!-- ``` r -->
<!-- install.packages('nanosvgr', repos = c('https://coolbutuseless.r-universe.dev', 'https://cloud.r-project.org')) -->
<!-- ``` -->

## `nsvg` object format

An `nsvg` object read from an SVG file using `nsvg_read()` is simply a data.frame
of geometric information and associated graphical parameters.

* Each row of the `nsvg` data.frame is one *shape* in the SVG
    * Each *shape* is made up of one-or-more *paths*
        * Each *path* is made up of one-or-more *cubic beziers*
            * Each *cubic bezier* is defined by 4 control points.

The columns in the `nsvg` data.frame are as follows:

* `shape_idx` - Shape index.  Each SVG is defined as a number of shapes, 
        with each shape having a number of paths
* `fill` - fill color
* `stroke` - stroke color
* `alpha` - opacity in range [0, 1]
* `lwd` - line width
* `linejoin`- Line join style.  'bevel', 'mitre' or 'round'
* `lineend` - Line end style. 'round', 'butt', 'square'
* `linemitre` - Line mitre limit
* `linedash` - Raw list of line dash lengths for each shape
* `fill_rule` - 'evenodd' or 'winding'
* `fill_type` - 'flat', 'linear', 'radial', 'none', 'undef'
* `gradient` - Radial or linear gradient information for this shape
* `stroke_type` - 'flat', 'linear', 'radial', 'none', 'undef'
* `beziers` - A list of data.frames - one data.frame for each
        shape containing the coordinates of the bezier control points 
        (Note: there are 4 control points for each cubic bezier).
    * `path_idx` - Index of path within shape
    * `bez_idx` - Index of bezier with path
    * `closed` - Is the path closed?
    * `x` - x coordinate of bezier control points
    * `y` - y coordinate of bezier control points
* `lty` - Line type. Either 'solid' or a string of up to 8 
        characters (from c(1:9, "A":"F")) may be given, giving the length of 
        line segments which are alternatively drawn and skipped.  See
        `?graphics::par` for deails on Line Type Specification
* `points` - A list of data.frames - one data.frame for each
        shape containing the polylines derived from the beziers. Data is the 
        same as `beziers` except the coordinates are for the polylines 
        derived from the beziers


## Reading an SVG

```{r out.height=500, out.width=500, fig.height=13, fig.width=13, fig.asp=1}
library(nanosvgr)
library(tibble)

nsvg <- nsvg_read("man/figures/tiger.svg")
class(nsvg)
# An 'nsvg' object consists of 
#   * raw geometry for the shapes
#   * graphical parameters e.g. fill color, stroke color etc
names(nsvg)
# The geometry data consists of a number of shapes.
# Each shape consists of 1-or-more paths 
# Each path may be open or closed (i.e. a line or a polygon)
# Each path is defined by one or more beziers
print(nsvg, width = Inf)
# The beziers for each path are defined by 4 x-coordinates and 4 y-coordinates.
# THe beziers for each path are collected in a list of data.frames in
# the column "beziers"
nsvg$beziers[[1]]
```

## Plotting with `nsvg`


```{r out.height=500, out.width=500, fig.height=13, fig.width=13, fig.asp=1}
nsvg <- nsvg_read("man/figures/tiger.svg")
grid::grid.newpage(); plot(nsvg)
```


```{r out.height=500, out.width=500, fig.height=13, fig.width=13, fig.asp=1}
nsvg <- nsvg_read("inst/gloves.svg") |> nsvg_invert_closedness()
grid::grid.newpage(); plot(nsvg)
```


## Example of geometry manipulation

As we now have access to the geometry of the SVG we can manipulate it
prior to plotting.

```{r}
nsvg <- nsvg_read("man/figures/tiger.svg") |> nsvg_scale(0.3)
nsvg <- nsvg_explode(nsvg, scale = 1.75)
grid::grid.newpage(); plot(nsvg)
```





## Plotting with grid



```{r out.height=500, out.width=500, fig.height=13, fig.width=13, fig.asp=1}
library(nanosvgr)
library(grid)

nsvg   <-  nsvg_read("man/figures/tiger.svg")
points <- nsvg_unnest_points(nsvg, gpars = TRUE)

grid::grid.newpage()
dfs <- split(points, points$shape_idx)
for (i in seq_along(dfs)) {
  df <- dfs[[i]]
  if (df$closed[1]) {
    grid.polygon(df$x, df$y, 
                 default.units = 'points', 
                 gp = gpar(col = nsvg$stroke[i], fill = nsvg$fill[i]))
  } else {
    grid.polyline(df$x, df$y, 
                 default.units = 'points', 
                 gp = gpar(col = nsvg$stroke[i], fill = nsvg$fill[i]))
  }
}
```

## Plotting with ggplot

```{r}
library(nanosvgr)
library(ggplot2)

nsvg   <-  nsvg_read("man/figures/tiger.svg")
points <- nsvg_unnest_points(nsvg, gpars = TRUE)
dfs    <- split(points, points$shape_idx)

p <- ggplot(points)

for (i in seq_along(dfs)) {
  df <- dfs[[i]]
  if (df$closed[1]) {
    p <- p + geom_polygon(data = df, aes(x, y, group = shape_idx, fill = I(fill), col = I(stroke), linewidth = I(lwd)))

  } else {
    p <- p + geom_path(data = df, aes(x, y, group = shape_idx, col = I(stroke)))
  }
}

p + 
  theme_bw() + 
  coord_equal()
```


## Plotting raw beziers with `{ggforce}`

```{r}
library(nanosvgr)
library(ggplot2)
library(ggforce)

nsvg    <- nsvg_read("man/figures/tiger.svg")
beziers <- nsvg_unnest_beziers(nsvg)

ggplot(beziers) +
  geom_bezier(aes(x, y, group = interaction(shape_idx, path_idx, bez_idx), color = as.factor(shape_idx))) +
  labs(
    title = "ggplot2 + ggforce::geom_bezier()"
  ) +
  theme_bw() +
  coord_equal() +
  theme(legend.position = 'none')
```


## Plotting raw beziers with `grid.bezier()`

```{r out.height=500, out.width=500, fig.height=13, fig.width=13, fig.asp=1}
library(nanosvgr)
library(grid)

nsvg    <- nsvg_read("man/figures/tiger.svg")
beziers <- nsvg_unnest_beziers(nsvg)
beziers <- split(beziers, with(beziers, interaction(shape_idx, path_idx, bez_idx, drop = TRUE)))

grid.newpage()
for (bezier in beziers) {
  grid.bezier(bezier$x, bezier$y, default.units = 'points')
}
```





